\documentclass[numsec,webpdf,modern,medium,namedate]{oup-authoring-template}

\onecolumn
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs,graphicx}
\usepackage{microtype}
\graphicspath{{./}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newcommand{\ind}{\mathbf 1}

\begin{document}
\journaltitle{Journal of the Royal Statistical Society Series B: Statistical Methodology}
\DOI{}
\copyrightyear{2026}
\pubyear{2026}
\access{}
\appnotes{Original Article}
\firstpage{1}
\title[Inference for hard classification likelihoods]{Inference for Hard Classification Likelihoods: Moving-Boundary Curvature and Feasible Uncertainty Quantification}

\author[1,$\ast$]{Samyajoy Pal\ORCID{0000-0003-1339-0979}}
\author[1]{Rouven E. Haschka}
\authormark{Pal and Haschka}
\address[1]{\orgdiv{Chair of Data Analytics}, \orgname{Technical University of Kaiserslautern-Landau}, \orgaddress{\street{Gottlieb-Daimler-Stra\ss e 42}, \postcode{67663}, \state{Kaiserslautern}, \country{Germany}}}
\corresp[$\ast$]{Samyajoy Pal, Chair of Data Analytics, Technical University of Kaiserslautern-Landau, 67663 Kaiserslautern, Germany. \href{mailto:samyajoy.pal@rptu.de}{samyajoy.pal@rptu.de}}

\abstract{Hard classification replaces posterior membership probabilities by
deterministic assignments and can make heterogeneous mixture fitting markedly
simpler.  It also creates a nonsmooth criterion whose classification regions
move with the parameter.  Standard errors computed after freezing the fitted
labels ignore that motion.  We study classification maximum-likelihood
estimators as potentially misspecified M-estimators, with the population
classification functional---rather than the ordinary mixture parameter---as
their explicit target, and focus on a selected local optimum.
For regular decision hypersurfaces, we show that the
population Hessian contains a positive semidefinite surface integral in
addition to the usual within-class Hessian.  Because this term reduces the
positive curvature, fixed-label sandwich inference generally understates
uncertainty.  We establish consistency, a local stochastic expansion,
asymptotic normality, local-bootstrap validity, and consistency of a feasible
kernel estimator of the surface term.  An eigenvalue-constrained
stabilization is asymptotically inactive but prevents nonpositive estimated
curvature in finite samples.  Across 31,000 primary simulation attempts involving
heterogeneous Normal--Student-$t$ and Normal--skew-normal mixtures, the
correction materially improves standard-error calibration and confidence
interval coverage under overlap and imbalance, while becoming small under
separation.  Two Dry Bean analyses show both successful local calibration and
the distinct instability caused by optimization-basin selection; diagnostics
identify basin instability.  The method therefore supplies feasible local
inference together with diagnostics that identify when a local Gaussian
summary is inadequate.}
\keywords{classification likelihood; decision boundary; finite mixture; M-estimation; nonsmooth inference; sandwich covariance}
\maketitle

\section{Introduction}
\label{sec:introduction}

Many estimators combine smooth component criteria through a parameter-dependent
maximum.  Examples arise whenever observations are assigned to the currently
best-fitting group and the parameters are then estimated using those hard
assignments.  Such procedures are attractive computationally: conditional on
the assignments, a large optimization can separate into simpler component
problems.  Statistically, however, the resulting objective is only piecewise
smooth.  A small parameter perturbation moves the interfaces between the
regions, so differentiating the criterion as though its fitted partition were
fixed omits part of its curvature.
The paper's main message is that hard classification creates moving decision
boundaries, and these boundaries contribute first-order curvature even though
they have probability zero.  Mixtures provide the main implementation, but
the theory applies more generally to max-type classification likelihoods built
from smooth component criteria.

Classification expectation--maximization (CEM) provides an important instance
of this problem.  Finite mixtures are standard tools for density estimation,
clustering, and latent-population analysis \citep{mclachlan2000finite}.  The EM
algorithm maximizes the observed-data likelihood through probabilistic
memberships \citep{dempster1977maximum}; CEM instead inserts a classification
step and assigns every observation to its largest weighted component density
\citep{celeux1992classification}.  Once labels are fixed, component updates
are partly decoupled.  This is especially useful in heterogeneous mixtures,
where different components use different distribution families and some
updates require numerical optimization \citep{pal2026flexible}.

Model-based clustering has developed a broad theory of mixture fitting,
identifiability, parsimonious covariance modelling, initialization, and order
selection \citep{titterington1985statistical,redner1984mixture,
banfield1993model,fraley2002model,keribin2000consistent}. Mixture likelihoods
can be singular or multimodal, motivating constrained estimation and careful
optimization diagnostics \citep{hathaway1985constrained,chen1995optimal,
mclachlan2008em}. Partition assessment and selection add a complementary
literature based on adjusted agreement, silhouettes, gaps, entropy, and
resampling stability \citep{hubert1985comparing,rousseeuw1987silhouettes,
tibshirani2001gap,celeux1996entropy,benhur2002stability,fang2012selection}.
These concerns are relevant to our basin diagnostics, but are distinct from
estimating the local curvature of a prespecified classification-likelihood
target.

Classification likelihood predates CEM. Likelihood-ratio clustering was
studied by \citet{scott1971clustering}, and
\citet{bryant1978classification} derived early large-sample properties of
classification maximum-likelihood estimates. Later work used classification
criteria for mixture-based clustering and model selection, most notably the
integrated completed likelihood \citep{biernacki2000icl} and conditional
classification likelihood \citep{baudry2015conditional}. These contributions
clarify that classification is a statistical criterion in its own right, but
they do not provide a general feasible sandwich covariance that estimates the
curvature contributed by moving multivariate decision faces in heterogeneous
component models.

Comparisons of mixture and classification sampling schemes likewise show that
hard classification targets must not be interpreted automatically as
ordinary-mixture parameters \citep{ganesalingam1989classification}. Related
order-selection work shows that nontrivial classification uncertainty changes
the asymptotic behaviour of complete-likelihood criteria \citep{hui2015order}.
For distance-based hard clustering, early distributional and
significance-test analyses further emphasize the role of a well-separated
population optimum \citep{hartigan1978asymptotic,bock1985probabilistic}.

The computational simplification changes both the estimand and the inference.
The CEM output maximizes a classification likelihood, not the ordinary
observed-data mixture likelihood.  Its natural population target is therefore
the maximizer of the expected classification criterion, and need not equal the
parameter that generated an ordinary finite mixture.  Moreover, the common
practice of conditioning on the fitted labels treats a data-dependent
partition as fixed.  This discards the effect of observations that cross a
decision boundary under an arbitrarily small parameter perturbation.  The
effect is negligible only when the data-generating density is sufficiently
small on the active boundaries; it can be first-order under component overlap.

The closest geometric precedent is the asymptotic theory of $k$-means.
\citet{pollard1981strong} established consistency and
\citet{pollard1982clt} derived a central limit theorem in which the geometry of
moving Voronoi cells is fundamental. Our setting replaces squared-distance
cells by general weighted component log-density comparisons, permits different
families and misspecification, identifies the active-face contribution in a
coordinate-invariant surface form, and develops a sample-based estimator of
that contribution suitable for routine sandwich inference.

This phenomenon is easy to miss because a regular boundary has probability
zero.  Probability zero eliminates a boundary contribution to the first
derivative, but not to the second.  We show that differentiation of the
population maximum produces a surface integral over each active decision
hypersurface.  Its integrand combines the data density on the boundary, the
sensitivity of the pairwise score contrast to the parameter, and the spatial
gradient transverse to the boundary.  The term is positive semidefinite in the
population Hessian.  It consequently subtracts from the usual positive
fixed-class curvature and generally increases the sandwich variance.

Our first contribution is a theory of the classification maximum-likelihood
functional under possible model misspecification.  We establish existence and
consistency, derive the moving-boundary Hessian, and prove a uniform local
empirical-process expansion for the nonsmooth maximum criterion.  These
results yield root-$n$ consistency, asymptotic linearity, and normality for a
sufficiently accurate local maximizer.  We also state explicit conditions
under which an algorithmic stationary solution inherits that distribution and
under which a nonparametric bootstrap is valid.  This separation is essential:
asymptotic theory for a population target does not guarantee that a multistart
algorithm or every bootstrap refit reaches its basin.

Our second contribution is feasible inference without estimating the full
data-generating density or numerically integrating over a hypersurface.  By a
coarea identity, the surface curvature is equivalently a Dirac-delta moment of
the pairwise score contrasts.  We estimate it using a one-dimensional kernel
around each active contrast, obtaining a procedure whose observation-level
cost is linear in sample size and does not require a grid in the observation
space.  We derive its bias, variance, oracle limit distribution, plug-in rate,
and consistency.  Because sampling variability can make the corrected
curvature nonpositive in finite samples, we introduce an eigenvalue-constrained
stabilization and prove that it is asymptotically inactive.
The construction uses the coarea formula \citep{federer1969geometric} and
standard one-dimensional kernel principles \citep{silverman1986density,
wand1995kernel}; its plug-in analysis is framed as a stochastic-equicontinuity
problem rather than a full multivariate density-estimation problem
\citep{andrews1994empirical}.

Our third contribution is an empirical assessment designed around the actual
statistical target and the method's operational limitations.  The simulation
study uses independently approximated population classification targets rather
than the ordinary mixture-generating parameters, retains failed fits, varies
overlap, imbalance, dimension, number of components, and heterogeneous family,
and examines bandwidth sensitivity.  Across the main designs, fixed-label
inference is most anti-conservative when the boundary carries appreciable
mass; the proposed correction moves estimated standard errors and coverage
toward their empirical values.  Additional large-sample runs distinguish slow
convergence from persistent failure.  In two real-data analyses we compare the
analytic covariance with a local nonparametric bootstrap and separately use
unrestricted bootstrap refits to diagnose basin-selection instability.

The contribution differs from ordinary likelihood information calculations,
including missing-information decompositions such as \citet{Louis1982FindingTO}.
Here the issue is not integration over an unobserved label conditional on a
smooth likelihood.  It is the motion of deterministic classification regions
inside a max-type population criterion.  It also differs from applying a
generic sandwich formula after differentiation: the relevant derivative
matrix itself contains a singular surface contribution.  The results connect
classical misspecified M-estimation \citep{Huber1967,white1982maximum} with the
geometry of hard classification.

Technically, max-type criteria sit within the wider literature on nonsmooth
M-estimation and empirical processes. Depending on how sharply the population
criterion changes, nonsmooth estimators can retain a Gaussian root-$n$ limit
or exhibit cube-root behaviour \citep{kim1990cube,sherman1993limiting}; the
regular active-face and positive-curvature assumptions below identify the
former regime. Our bootstrap statement builds on classical nonparametric
bootstrap theory \citep{efron1979bootstrap,efron1993introduction}, while the
application diagnostics recognize that clustering instability is a separate
inferential object.

\subsection{A one-dimensional illustration of the boundary term}
\label{sec:one-dimensional-illustration}
Before introducing the general notation, a scalar example makes the source
and magnitude of the additional curvature transparent.  
Consider two components on the real line and suppose that their contrast
$g(x;\boldsymbol{\theta})=q_1(x;\boldsymbol{\theta})-
q_2(x;\boldsymbol{\theta})$ has one active regular boundary point $x_0$ at
$\boldsymbol{\theta}^*$.  The surface correction in
Theorem~\ref{thm:population-curvature} then reduces to
\begin{equation}
 \mathbf B=\frac{p_0(x_0)\nabla_{\boldsymbol\theta}g(x_0;\boldsymbol\theta^*)
 \nabla_{\boldsymbol\theta}g(x_0;\boldsymbol\theta^*)^\mathsf T}
 {|\nabla_x g(x_0;\boldsymbol\theta^*)|}.
 \label{eq:one-dimensional-boundary}
\end{equation}
Thus the correction is large when the data density is high at the decision
boundary, all else equal.  It is small when the boundary lies in a low-density
region, which explains why fixed-label inference can be adequate for strongly
separated groups.

\newpage
The general development below replaces the single point $x_0$ by active
hypersurfaces and turns this population identity into feasible inference.
Section~\ref{sec:methodology} defines the estimand, derives the active-boundary
curvature, establishes the limit theory, and constructs the feasible
covariance estimator. Section~\ref{sec:results} reports the simulation study
and real-data analyses. Sections~\ref{sec:discussion} and the Conclusion state
the implications and limitations. Technical proofs, complete empirical
summaries, direct Gaussian-limit diagnostics, failure accounting, and
application-level parameter tables are provided in the supplement.

\section{Methodology}
\label{sec:methodology}

\subsection{Classification likelihood and the population target}
\label{sec:framework}
\input{sections/framework}

\subsection{Local geometry of active decision boundaries}
\label{sec:geometry}
\input{sections/geometry}

\subsection{Asymptotic theory}
\label{sec:theory}
\input{sections/consistency}
\input{sections/asymptotics}

\subsection{Feasible boundary-curvature estimation}
\label{sec:estimation}
\input{sections/feasible_inference}

\subsection{Computation and implementation}
\label{sec:computation}
\input{sections/computation}

\section{Results}
\label{sec:results}

\subsection{Simulation study}
\label{sec:simulations}
\input{sections/simulation_results}

\subsection{Real data analysis}
\label{sec:applications}
\input{sections/applications}

\section{Discussion}
\label{sec:discussion}
Hard classification creates a distinctive inferential problem: fitted labels
can be locally constant for almost every individual observation even though
their population partition changes at first order.  The missing curvature is
concentrated on decision hypersurfaces.  Our surface representation makes this
mechanism explicit, and its positive semidefinite sign explains why treating
the labels as fixed tends to produce standard errors that are too small.

The feasible estimator turns that geometric identity into a practical
procedure.  It smooths only a scalar log-score contrast, regardless of the
dimension of the observations, and it samples the unknown data distribution
directly.  This is preferable to integrating a fitted mixture density over an
estimated surface, which would impose correct specification precisely where
the theory does not require it.  The simulations and the Dermason--Seker
analysis show that the correction can improve local calibration without the
cost of a large bootstrap.  Under clear separation the correction contracts,
as predicted by the surface formula.

\paragraph{What the method does not do.}
The method does not target the ordinary mixture parameter: its estimand is the
population hard-classification optimum.  It does not solve uncertainty from
label selection or component-family selection.  Nor does it guarantee valid
inference near nonregular configurations with multiple active boundaries.
Regular active faces,
a unique local target, and positive population curvature are substantive
conditions.  Nearly tangent boundaries, appreciable multiple ties, vanishing
components, or weakly identified skewness can invalidate the approximation or
slow its convergence.  Global family assignment and optimization-basin
selection are discrete sources of uncertainty.  The unrestricted bootstrap
results, particularly for Cali--Barbunya, show that they should not be hidden
inside a Wald standard error.
The method estimates local sampling uncertainty conditional on the selected
basin; it does not account for global optimization-basin selection.  When
unrestricted bootstrap refits often move to another basin, corrected Wald
intervals should be reported only as conditional local summaries.

These qualifications suggest a reporting protocol.  Alongside corrected
standard errors, an analysis should provide component weights, boundary kernel
support, the raw relative boundary eigenvalue, any stabilization factor,
multistart objective gaps, and the fraction of local bootstrap fits that pass
their convergence diagnostics.  When unrestricted refits frequently change
basin, conditional local intervals remain interpretable for the selected
solution but should not be presented as a complete uncertainty statement.

Several extensions are promising.  Uniform inference as two population maxima
approach one another would require a nonregular, set-valued limit theory.
Joint selection and estimation could instead use selective or subsampling
methods.  Higher-order or locally adaptive bandwidths may improve estimation
when boundary density varies sharply, while automatic differentiation could
replace finite differences in large parameter spaces.  More broadly, the same
surface-curvature principle applies to other estimators built from maxima of
smooth criteria.  Developing this connection beyond mixture classification is
an important direction for future work.

\section{Conclusion}
The inferential consequence of hard classification is geometric: decision
boundaries move with the fitted parameter and contribute first-order
curvature even though the boundaries themselves have probability zero. The
proposed surface correction estimates this missing curvature with a
one-dimensional kernel calculation and yields materially better calibrated
local uncertainty where fixed-label inference is most anti-conservative. The
theory, simulations, and applications also delimit the claim: corrected Wald
inference describes a regular local optimum and cannot absorb instability from
competing optimization basins or nearly nonregular classification faces.
Reporting the correction together with boundary support, stabilization, and
basin diagnostics therefore supplies both an inferential method and an
explicit warning when a Gaussian local summary is insufficient.

\section*{Data and code availability}
The Dry Bean data are publicly available from the UCI Machine Learning
Repository under a Creative Commons Attribution 4.0 licence
\citep{uci2020drybean}. Code, prespecified screening output, raw job manifests,
and scripts producing every table and figure are maintained at
\url{https://github.com/samyajoypal/CEM_Asymptotics}.

\section*{Conflict of interest}
The authors declare that they have no conflict of interest.

\section*{Funding}
No external funding was received for this study.

\section*{Author contributions}
S.P. contributed conceptualization, methodology, software, formal analysis,
visualization, and preparation of the original draft. R.E.H. contributed
review and editing of the manuscript. Both authors approved the final
manuscript.
\bibliographystyle{abbrvnat}
\bibliography{references}
\end{document}
